---
title: "Data Visualisation"
author: "Eugene"
date: "`r format(Sys.Date(), '%B %d %Y')`"
output:
  xaringan::moon_reader:
    css: [TUD_workshop, tamu-fonts]
---

class: center, inverse

# Workshop - dplyr

### Perform the Following Tasks:


__*Problem 1*__

Take the _us_contagious_diseases_ dataset from the _dslabs_ library. _filter()_ the dataset for the disease _Smallpox_ in the state of _Wisconsin_. This should give you 25 rows. Use this to produce the plot below.

```{r filter, echo = F, message=F, warning=F, fig.height=5}
library(tidyverse)
library(dslabs)
us_contagious_diseases %>% 
  filter(disease == "Smallpox", state == "Wisconsin") %>% 
  ggplot(aes(year, count)) + 
  geom_line(col = "springgreen4") + 
  geom_point(col = "darkslategrey")
```

---

class: center, inverse

__*Problem 2*__

Take the _research_funding_rates_ dataset from the _dslabs_ library. Use _select()_ to print out the dataset with only the columns _discipline_, _success_rates_men_, and _success_rates_women_. The dataframe should look as below:

```{r select_example, echo = F, message=F, warning=F}
library(knitr)
library(kableExtra)

research_funding_rates %>% 
  select(discipline, success_rates_men, success_rates_women) %>% 
  kable() %>% 
  kable_styling(stripe_color = "palegreen2",
                bootstrap_options = "striped")
```

---

class: center, inverse

__*Problem 3*__

Take the table from problem 2 and use _arrange(desc())_ to modify it so that the rows are ordered by decreasing values of _applications_total_

```{r arrange_example, echo = F, message=F, warning=F}

research_funding_rates %>%   
  arrange(desc(applications_total)) %>% 
  select(discipline, success_rates_men, success_rates_women)  %>% 
  kable() %>% 
  kable_styling(stripe_color = "palegreen2",
                bootstrap_options = "striped")
```

---

class: center, inverse

__*Problem 4*__

Again, take the _research_funding_rates_ dataset from the _dslabs_ library. Make a new column using _mutate()_ that shows the overall success rate for each discipline. Print out the dataset with only the columns _discipline_ and _success_rates_, ordered by success rate. The dataframe should look as below:

```{r mutate_example, echo = F, message=F, warning=F}

research_funding_rates %>% 
  mutate(success_rate = awards_total / applications_total) %>% 
  select(discipline, success_rate) %>% 
  mutate(success_rate = success_rate %>% round(3)) %>% 
  arrange(desc(success_rate)) %>% 
  kable() %>% 
  kable_styling(stripe_color = "palegreen2",
                bootstrap_options = "striped")
```

---

class: center, inverse

__*Problem 5*__

Take the _polls_us_election_2016_ dataset from the _dslabs_ library. _group_by()_ the _grade_ column and _summarise()_ to calculate the average sample size for each grade. _arrange(desc())_ the table in decreasing _average_sample_size_. The dataframe should look as below:

```{r group_example, echo = F, message=F, warning=F}
polls_us_election_2016 %>% 
  group_by(grade) %>% 
  summarise(average_sample_size = mean(samplesize, na.rm = T) %>% round(0)) %>% 
  arrange(desc(grade)) %>% drop_na() %>% 
  kable() %>% 
  kable_styling(stripe_color = "palegreen2",
                bootstrap_options = "striped")

```

Take the output from this dataframe and make a barchart
---

class: center, inverse

__*Problem 6*__


Take the _msleep_ dataset from the _ggplot2_ library. Get rid of incomplete rows (use _msleep %>% drop_na()_). Use _select(-c())_ to get rid of the _genus_, _vore_, _order_, _conservation_, _brainwt_, and _bodywt_ columns. Make this in to a long form data frame with columns, _name_, _type_of_sleep_, and _hours_ using the _pivot_longer()_ command setting _col_ to _-name_, _names_to_ to _type_of_sleep_ , and _values_to_ to _hours_. The dataframe should look as below:


```{r pivot_example, echo = F, message=F, warning=F}
msleep %>% 
  drop_na() %>% 
  select(-c("genus", "vore", "order", "conservation", "brainwt", bodywt)) %>% 
  arrange(desc(sleep_total)) %>% 
  head(3) %>% 
  pivot_longer(cols = -name,
               names_to = "type_of_sleep", 
               values_to = "hours") %>% 
  mutate(hours = hours %>% signif(2)) %>% 
  kable() %>% 
  kable_styling(stripe_color = "palegreen2",
                bootstrap_options = "striped")

```



- 
